---
title: "Kiriakos_MSDS 660_X70_Week 1 Basic Concepts"
author: "Cathy Kiriakos"
date: "October 29, 2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Week 1 Basic Concepts using: 
Introduction to R (pp. 41-52) James, G., Witten, D., Hastie, T., and Tibshirani, R. (2013). An Introduction to Statistical Learning with Applications in R. Springer. Retrieved from: http://www-bcf.usc.edu/~gareth/ISL/ISLR%20Fourth%20Printing.pdf

# Purpose: 
The purpose of this analysis is to provide a basic understanding of the foundational principles of Statistical Methods in this case using R.  We will explore data provided by the Federal Reserve Bank of St. Louis (FRED).  As it relates to my field of work we will explore the effective Fed Funds Rate: 

Below is a summary of the data from the Saint Louis Fed: 
Source: Board of Governors of the Federal Reserve System (US)  Release: H.15 Selected Interest Rates  
Units:  Percent, Not Seasonally Adjusted

Frequency:  Monthly

Averages of daily figures.

The federal funds rate is the interest rate at which depository institutions trade federal funds (balances held at Federal Reserve Banks) with each other overnight. When a depository institution has surplus balances in its reserve account, it lends to other banks in need of larger balances. In simpler terms, a bank with excess cash, which is often referred to as liquidity, will lend to another bank that needs to quickly raise liquidity. (1) The rate that the borrowing institution pays to the lending institution is determined between the two banks; the weighted average rate for all of these types of negotiations is called the effective federal funds rate.(2) The effective federal funds rate is essentially determined by the market but is influenced by the Federal Reserve through open market operations to reach the federal funds rate target.(2)
The Federal Open Market Committee (FOMC) meets eight times a year to determine the federal funds target rate. As previously stated, this rate influences the effective federal funds rate through open market operations or by buying and selling of government bonds (government debt).(2) More specifically, the Federal Reserve decreases liquidity by selling government bonds, thereby raising the federal funds rate because banks have less liquidity to trade with other banks. Similarly, the Federal Reserve can increase liquidity by buying government bonds, decreasing the federal funds rate because banks have excess liquidity for trade. Whether the Federal Reserve wants to buy or sell bonds depends on the state of the economy. If the FOMC believes the economy is growing too fast and inflation pressures are inconsistent with the dual mandate of the Federal Reserve, the Committee may set a higher federal funds rate target to temper economic activity. In the opposing scenario, the FOMC may set a lower federal funds rate target to spur greater economic activity. Therefore, the FOMC must observe the current state of the economy to determine the best course of monetary policy that will maximize economic growth while adhering to the dual mandate set forth by Congress. In making its monetary policy decisions, the FOMC considers a wealth of economic data, such as: trends in prices and wages, employment, consumer spending and income, business investments, and foreign exchange markets.
The federal funds rate is the central interest rate in the U.S. financial market. It influences other interest rates such as the prime rate, which is the rate banks charge their customers with higher credit ratings. Additionally, the federal funds rate indirectly influences longer- term interest rates such as mortgages, loans, and savings, all of which are very important to consumer wealth and confidence.(2)
References
(1) Federal Reserve Bank of New York. "Federal funds." Fedpoints, August 2007.
(2) Board of Governors of the Federal Reserve System. "Monetary Policy".

Citation:
Board of Governors of the Federal Reserve System (US), Effective Federal Funds Rate [FEDFUNDS], retrieved from FRED, Federal Reserve Bank of St. Louis; https://fred.stlouisfed.org/series/FEDFUNDS, October 28, 2019.

https://fred.stlouisfed.org/series/FEDFUNDS

We will start by importing the dataset: 
```{r, import.dataset}
fedFunds <- read.csv("FEDFUNDS.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)
```
Calling head to get a view of the data: 
```{r, view}
head(fedFunds)
```
By looking at the head of the data, we can see that the date was formatted as Character, we will format the column to be a date with the call below: 
```{r,formatdate}
fedFunds$DATE<- as.Date(fedFunds$DATE,format = "%m/%d/%Y")
```
Again with a quick view we can see that it is now in a date format
```{r,viewtwo}
head(fedFunds)
```
Now that the date is formatted appropriately, we will address the fedfunds 
```{r,formatrate}
fedFunds$FEDFUNDS<-as.numeric(fedFunds$FEDFUNDS)
```

```{r,viewthree}
head(fedFunds)
```

In the call below we will get a summary of the data set, and can see that it begins in 1954 and ends in 2019, the the minimum rate observed was 0.070, the maximum is 19.10, the mean is 4.785, there are 783 observations
```{r,summary}
summary(fedFunds)
```
Now we will get a look at the data  by calling histogram: 
```{r,hist}
hist(fedFunds$FEDFUNDS, main = "Histogram of Fed Funds Rate 1954-2019",xlab = "Fed Funds Rate", ylab = "Interest Rate")
```
Now we will create a line chart
```{r,plot}
plot(fedFunds$DATE,fedFunds$FEDFUNDS,type = "s", col = "blue", xlab = "Date", ylab = "Fed Funds Rate", main = "Fed Funds Timeseries 1954-2019")
```


