---
title: "Kiriakos_MSDS660_X70_Week5_TwoWayANOVA"
author: "Cathy Kiriakos"
date: "November 25, 2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
"
## MSDS 660_X70 Week 5 Two-Way ANOVA
The purpose of this assignment is to analyze income data between male and female respondents using two-way ANOVA.

A researcher wants to investigate the income by Region (North, South, East, West) and gender (male, female). A sample of 24 people combining region and gender are examined. The Income ($1000USD) data are shown as:

First we must create the data frame: 
```{r,dfcreate}
region <- c("North", "North", "North", "South", "South", "South", "East", "East", "East", "West", "West", "West","North", "North", "North", "South", "South", "South", "East", "East", "East", "West", "West", "West")
Income <- c(50,60,45,40,53,68,92,74,87,120,150,85,45,72,65,55,75,48,87,105,79,95,78,115)
Sex <- c("male","male","male","male","male","male","male","male","male","male","male","male","female","female","female","female","female","female","female","female","female","female","female","female")
IncomeData<-data.frame("Region" = region, "Sex" = Sex, "Income" = Income)
summary(IncomeData)
```
First we will test our data using ggpairs for normality: 
```{r,testdata, eval=FALSE}
ggpairs(IncomeData,lower=list(combo=wrap("facethist", binwidth=0.8)))
```
Ok doing our testing we can see that we have normal distributions, We can now run our levene test for homogeneity of variance: 
```{r,leveneMale}
library(car)
leveneTest(Income ~ Sex*Region, data = IncomeData)
```
We can see that our P Value is greater than 0.05, telling us that we can not reject the null hypothesis that the variance is equal across groups, suggesting that conducting a non-parametric test might be a better option.  
```{r,leveneFemale}
leveneTest(Income ~ Sex, data = IncomeData)
```
Now we can get a view of our two-way ANOVA analyzing how income is impacted by sex in region using our data set:
```{r,residplot}
aov.income <- aov(Income ~ Sex + region, data = IncomeData)
summary(aov.income)
```
With our output table we can see that while sex and sex*region has a high P-value where we will reject our null hypothesis that the group means are not statistically different, but region at 99% confidence has statistically different group means. 

Now we can compute the two-way ANOVA using the interaction effect of sex and region: 
```{r,anovainteraction}
aov.income2<- aov(Income ~ Sex*region + Sex + region, data = IncomeData)
summary(aov.income2)
```
Again, we can see that even the combined sex/region coefficient is not statistically significant, but region is.  We can get a quick visual of the interaction effect below: 
The visualization script has been obtained using Cran' SJ Plot Package information from the link below:
https://cran.r-project.org/web/packages/sjPlot/vignettes/plot_interactions.html
```{r, plotinteraction}
library(sjPlot)
library(sjmisc)
library(ggplot2)
data(IncomeData)
theme_set(theme_sjplot())

IncomeData$Sex <- to_factor(IncomeData$Sex)

#fit model with interaction: 
fit <- lm(Income ~ Sex*Region + Sex + region, data = IncomeData)
plot_model(fit, type = "pred", terms = c("Sex", "region"))
```

```{r}
plot_model(fit, type = "int")
```
As anticipated because we saw the interaction effects in our ANOVA, the Sex Region interaction effect is not impactful in our model. 


W can see that Region is the only coefficient where the means are statistically significant at 99% confidence interval, whereas sex plays a less impactful role in in our independent variable income. 

